---
title: "Assignment 1"
author: "Suman"
date: "5/18/2020"
output:
  pdf_document: default
  html_document: default
---

## Loading and preprocessing the data
```{r, echo=TRUE}
library(knitr)
opts_chunk$set(echo =TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
setwd("~/R/Reproducibel_Research")
data <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character",
                                                                  "integer"))
data$date <- ymd(data$date)
```

## What is mean total number of steps taken per day?

1. Calculating the steps taken

```{r , echo=TRUE}
steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```

2. Constructing the plot using ggplot
```{r}
ggplot(steps, aes(x = steps)) +
  geom_histogram(fill = "firebrick", binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```

3. Calculating the mean and the median

```{r}
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)
mean_steps
median_steps
```
## What is the average daily activity pattern?

1. Calcualting the acerage steps

```{r}
interval <- data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))
```
2. Plotting the graph
```{r}
ggplot(interval, aes(x=interval, y=steps)) +
  geom_line(color = "firebrick")
```
3. Finding the max steps
```{r}
interval[which.max(interval$steps),]
```

## Imputing missing values

1. Filling in the missing values
```{r}
data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
```

2. Calculating the steps
```{r}
steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```

3. Plotting the graph
```{r}
ggplot(steps_full, aes(x = steps)) +
  geom_histogram(fill = "firebrick", binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")
```

4. Calculating the median and the mean

```{r}
mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)
mean_steps_full
median_steps_full
```

## Are there differences in activity patterns between weekdays and weekends?

1. Whether weekday or weekend:
```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
```


2. average steps in 5 min intervals

```{r}
interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
```

3. Graph

```{r}
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```
